### 2009-11-13 src/Docs/Intro.tex
Reorganised Epitome - added introduction (ish) and moved perversity to an appendix

\section{In the beginning}

We really should right something informative here.