### 2009-12-06 src/Docs/Intro.tex
Recommended readings for the Desc universe.

\section{In the beginning}

We really should write something informative here.

\subsection{Recommended readings}

The type-checker is a bidirectional
one~\cite{turner:bidirectional_tc}. A remotely related type-checker
has been described in the context of ETT~\cite{chapman:ett}.

Normalization is achieved by evaluation~\cite{dybjer:nbe,
  dybjer:dependent_types_work}. The implementation has been described
in James Chapman's work~\cite{chapman:phd}. A graspable introduction
to both normalization by evaluation and bidirectional type-checking
\`a la Epigram can be found in Boutillier's
report~\cite{boutillier:report}.

The story for names has been told by McBride and
McKinna~\cite{mcbride:free_variable}. \emph{Ite messa est}.

For containers, there is a lot to say. So, I will not say anything for
the moment.

Concerning the |Desc| universe, Morris et al. \cite{morris:spf} wrote
a clear article, covering that topic and much more. In particular,
they show why and how |Box| and |mapBox| can automatically derived
from |Desc|. Conor's work on Ornamemts \cite{mcbride:ornaments} builds
up from a simplified |Desc| universe: some insights can be found there
too. Finally, the authoritative source shall be cited: Dybjer et
al. on induction-recursion \cite{dybjer:ir_axiom, dybjer:ir_algebra,
  dybjer:iir}.