### 2010-04-24 papers/containers-for-dummies/containers.tex
Containers for Dummies: first import, just infrastructure.

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{url}
\usepackage{xspace}
\usepackage{ifthen}

\newcommand{\DRAFT}{}

%% Structure
\ifthenelse{\isundefined{\DRAFT}}
           {\newenvironment{structure}
                           {\comment}
                           {\endcomment}}
           {\newenvironment{structure}
                           {\footnotesize\verbatim}
                           {\endverbatim}}

%% Written bits of Structure
\newenvironment{wstructure}{\comment}{\endcomment}

%% Comments
\ifthenelse{\isundefined{\DRAFT}}
           {\newcommand{\note}[1]{}}
           {\setlength{\marginparwidth}{0.7in}
            \newcommand{\note}[1]{\-\marginpar[\raggedright\footnotesize #1]%
                                 {\raggedright\footnotesize #1}}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Draft: Containers for Dummies}
\author{The MSP Doctoral Crew
        \footnote{Correspondent author: 
                  \texttt{dagand@cis.strath.ac.uk}}}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Content
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
The idea of writing this document arised during a MSP meeting, when I
(Pierre) was begging for some explanation of Michael Abbott's PhD
thesis~\cite{abbott:phd-thesis}. Neil answered that I should develop
my own understanding of the Container story instead. Hank acknowledged
that resources on Containers were rather technical and going in many
directions, mostly because each member of the Container Mafia have his
personal take on it.

It was suggested to write a document that would serve as an
introduction as well as a reference on Containers. Being an
introduction, this document should allow newcomers to become familiar
with our technology, thanks to a plethora of examples and
exercices. Being a reference, it should be an exhaustive overview of
the Container landscape, with clear definitions and theorems.
\end{abstract}

\begin{structure}
Some braindump:

* I have several pages of definitions from Hank. That could make a
  first skeleton.

* As suggested by Hank, we should first focus on the table of content,
  the content will follow.

* We will inevitably end up talking to the Programmer, the Type
Theorist and the Category Theorist. That would be good to organize the
document to allow people to build on their intuition and then move on
what they are less familiar with. We could (graphically) signal when
some paragraph goes into technicalities of a particular field.
\end{structure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrvnat}
\bibliography{containers}

\end{document}