### 2010-07-25 papers/icfp-2010-desc/paper_discussion.tex
Discussion rejigged

\section{Discussion}

In this paper, we have presented a universe of datatypes for a
dependent type theory. We started from an unremarkable type theory
with dependent functions and tuples, but relying on few other
assumptions, especially where propositional equality is concerned.  We
added finite enumeration sufficient to account for constructor choice,
and then we built coding systems, first (as a learning experience) for
simple ML-like inductive types, then for the indexed inductive
families which dependently typed programmers in Agda, Coq and Epigram
take for granted. We adopt a bidirectional type propagation mechanism
to conceal artefacts of the encoding, giving a familiar and practicable
constructor-based presentation to data.

Crucially to our approach, we ensure that the codes describing
datatypes inhabit a datatype with a code. In a stratified setting, we
avoid paradox by ensuring that this type of codes lives uniformly one
level above the types the codes describe. The adoption of ordinary
data to describe types admits datatype-generic operations implemented
just by ordinary programming. In working this way, we make
considerable use of type equality modulo open computation, silently
specialising the types of generic operations as far as the datatype
code for any given usage is known.

\subsection{Related work in Generic Programming}

\begin{wstructure}
!!! Need Help !!!
<- Comparison with Induction Recursion
    ???
\end{wstructure}

\begin{wstructure}
!!! Need Help !!!
<- Related Work
    <- Generic in simply-typed functional languages
        <- PolyP \cite{jansson:polyp}
        <- Generic Haskell \cite{hinze:generic-haskell}
        <- Scratch your boilerplate \cite{spj:syb}
\end{wstructure}

Generic programming is a vast topic. We refer our reader to
\citet{garcia:generic-comparative-study} for a broad overview of
generic programming in various languages. For Haskell alone, there is
a myriad of proposals: 
\citet{hinze:generic-approach-comparative} and
\citet{rodriguez:generic-libs-comparative} provide useful comparative
surveys.

Our approach follows the polytypic programming style, as initiated
by PolyP~\cite{jansson:polyp}. Indeed, we build generic functions by
induction on pattern functors, exploiting type-level computation
to avoid the preprocessing phase: our
datatypes are, natively, nothing but codes.

We have the \emph{type-indexed datatypes} of Generic
Haskell~\cite{hinze:generic-haskell} for free.
From one datatype, we can compute
others and equip them with relevant structure: the free monad
construction provides one example. Our approach to encoding datatypes
as data also sustains \emph{generic
  views}~\cite{holdermans:generic-view}, allowing us to rebias the
presentation of datatypes conveniently. Tagged descriptions,
giving us a sum-of-sigmas view, are a natural example.

Unlike Generic Haskell, we do not support polykinded
programming~\cite{hinze:polytypic-polykinded}. Our descriptions are
limited to endofunctors on $\Set^I$. Whilst indexing is known
to be sufficient to \emph{encode} a large class of higher-kinded
datatypes~\citep{DBLP:conf/ifip2-1/AltenkirchM02}, we should rather
hope to work in a more compositional style. We are free to write
higher-order programs manipulating codes, but is not yet clear whether
that is sufficient to deliver abstraction at higher kinds.
Similarly, it will be interesting to see whether
arity-generic programming~\cite{weirich:arity-generic} arises
just by computing with our codes, or whether a richer abstraction is
called for.

The Scrap Your Boilerplate~\cite{spj:syb} (SYB) approach to generic
programming offers a way to construct generic functions, based on
dynamic type-testing via the $\CN{Typeable}$
type class. SYB cannot compute types from codes, but its dynamic character
does allow a more flexible \emph{ad hoc} approach to generic data
traversal. By maintaining the correspondence between codes and types
whilst supporting arbitrary inspection of codes, we pursue the same
flexibility statically. The substitutive character of $\SYMBIDesc$
may allow us to observe and exploit \emph{ad hoc} substructural
relationships in data, but again, further work is needed if we
are to make a proper comparison.

\begin{wstructure}
    <- Generic in dependent types
        <- Norell \cite{norell:msc-thesis}
        <- Polytypic prog in Coq \cite{verbruggen:polytype-coq}
        <- Universes for generic prog \cite{benke:universe-generic-prog}
\end{wstructure}

\subsection{Generic Programming with Dependent Types}

Generic programming is not new to dependent
types. \citet{DBLP:conf/ifip2-1/AltenkirchM02} developed a universe
of polykinded types in Lego;
\citet{norell:msc-thesis} gave a formalization of
polytypic programming in Alfa, a precursor to Agda;
\citet{verbruggen:polytype-prog-coq, verbruggen:polytype-coq} provided
a framework for polytypic programming in the Coq theorem
prover. However, these works aim at \emph{modelling} PolyP or Generic
Haskell in a dependently-typed setting for the purpose of proving
correctness properties of Haskell code. Our approach is different in
that we aim at building a foundation for datatypes, in a
dependently-typed system, for a dependently-typed system.

Closer to us is the work of \citet{benke:universe-generic-prog}. This
seminal work introduced the usage of universes for developing generic
programs. Our universes share similarities to theirs: our universe of
descriptions is similar to their universe of iterated induction, and
our universe of indexed descriptions is equivalent to their universe
of finitary indexed induction. This is not surprising, as we share the
same source of inspiration, namely induction-recursion.

However, we feel ready to offer a more radical prospectus. Their
approach is generative: each universe extends the base type theory
with both type formers and elimination rules. Thanks to levitation, we
rely only on a generic induction and a specialised $\SYMBswitchD$,
closing the type theory. We explore \emph{programming} with codes, but
also how to conceal the encoding when writing `ordinary' programs.

\subsection{Metatheoretical Status}

The \(\Set:\Set\) approach we have taken in this paper is convenient
from an experimental perspective, and it has allowed us to focus
primarily on the encoding of universes, leaving the question
of stratification (and with it, consistency, totality,
and decidability of type checking) to one side. However, we must
surely face up to the latter, especially since we have taken up the
habit of constructing `the set of all sets'. A proper account requires
a concrete proposal for a system of stratified universes which allows
us to make `level-polymorphic' constructions, and we are actively
pursuing such a proposal. We hope soon to have something to prove.

In the meantime, we can gain some confidence by systematically
embedding predicative fragments of our theory within
systems which already offer a universe hierarchy. We can, at the
very least, confirm that in UTT-style theories with conventional
inductive families of types~\citep{luo:utt}, as found in Coq
(and in Agda if one avoids experimental extensions), we build the
tower of universes we propose, cut off at an arbitrary height.
It is correspondingly clear that some such system can be made to
work, or else that other, longer-standing tools are troubled.

A metatheoretical issue open at time of writing concerns the
size of the \emph{index} set \(\V{I}\) in \(\IDesc{\V{I}}\). Both Agda
and recent versions of Coq allow inductive families with \emph{large}
indices, effectively allowing `higher-kind' fixpoints on
\(\Set^{\Set}\) and more. They retain the safeguard that the types of
\emph{substructures} must be as small as the inductively
defined superstructure. This liberalisation allows us large index sets
in our models, but whilst it offers no obvious route to paradox by
smuggling a large universe inside a small type, it is not yet known
to be safe. We can restrict \(\V{I}\) as necessary to avoid
paradox, provided \(\Unit\), used to index \(\SYMBIDesc\) itself, is
`small'.

\subsection{Further Work}

Apart from the need to nail down a stratified version of the system and
its metatheory, we face plenty of further problems and opportunities.
Although we have certainly covered Luo's criteria for inductive
families~\cite{luo:utt}, there are several dimensions in which to
consider expanding our universe.

Firstly, we seek to encompass
\emph{inductive-recursive} datatype families~\cite{dybjer:iir}, allowing
us to interleave the definition and interpretation of data in intricate
and powerful ways. This interleaving seems particularly useful when
reflecting the syntax of dependent type systems.

Secondly, we should very much like to extend our
universe with a codes for internal fixpoints, as
in~\cite{DBLP:conf/types/MorrisAM04}. The external knot-tying
approach we have taken here makes types like `trees with lists of subtrees'
more trouble than they should be. Moreover, if we allow the alternation
of least and greatest fixpoints, we should expect to gain types which are
not readily encoded with one external \(\SYMBIMu\).

Thirdly, it would be fascinating to extend our universe with dedicated
support for syntax with \emph{binding}, not least because a universe
with internal fixpoints has such a syntax. Harper and Licata have
demonstrated the potential for and of such an
encoding~\citep{DBLP:conf/icfp/LicataH09}, boldly encoding the invalid
definitions along with the valid. A more conservative strategy might
be to offer improved support for datatypes indexed by an extensible
context of free variables, with the associated free monad structure
avoiding capture as shown by~\citet{altenkirch:monadic-lambda}.

Lastly, we must ask how our new presentation of datatypes
should affect the tools we use to build software. It is not enough to
change the game: we must enable better play. If datatypes are data,
what is design?