### 2011-02-16 notes/Epigram2/Epigram2.tex
NatFinVec in Kit

\documentclass{article}
\usepackage{palatino}
\usepackage{a4wide}
\usepackage{alltt}
\usepackage{bbold}

%%%%%%%%% INFERENCE RULES

\newlength{\rulevgap}
\setlength{\rulevgap}{0.05in}
\newlength{\ruleheight}
\newlength{\ruledepth}
\newsavebox{\rulebox}
\newlength{\GapLength}
\newcommand{\gap}[1]{\settowidth{\GapLength}{#1} \hspace*{\GapLength}}
\newcommand{\dotstep}[2]{\begin{tabular}[b]{@{}c@{}}
                         #1\\$\vdots$\\#2
                         \end{tabular}}
\newlength{\fracwid}
\newcommand{\dotfrac}[2]{\settowidth{\fracwid}{$\frac{#1}{#2}$}
                         \addtolength{\fracwid}{0.1in}
                         \begin{tabular}[b]{@{}c@{}}
                         $#1$\\
                         \parbox[c][0.02in][t]{\fracwid}{\dotfill} \\
                         $#2$\\
                         \end{tabular}}
\newcommand{\Rule}[2]{\savebox{\rulebox}[\width][b]                         %
                              {\( \frac{\raisebox{0in} {\( #1 \)}}       %
                                       {\raisebox{-0.03in}{\( #2 \)}} \)}   %
                      \settoheight{\ruleheight}{\usebox{\rulebox}}          %
                      \addtolength{\ruleheight}{\rulevgap}                  %
                      \settodepth{\ruledepth}{\usebox{\rulebox}}            %
                      \addtolength{\ruledepth}{\rulevgap}                   %
                      \raisebox{0in}[\ruleheight][\ruledepth]               %
                               {\usebox{\rulebox}}}
\newcommand{\Case}[2]{\savebox{\rulebox}[\width][b]                         %
                              {\( \dotfrac{\raisebox{0in} {\( #1 \)}}       %
                                       {\raisebox{-0.03in}{\( #2 \)}} \)}   %
                      \settoheight{\ruleheight}{\usebox{\rulebox}}          %
                      \addtolength{\ruleheight}{\rulevgap}                  %
                      \settodepth{\ruledepth}{\usebox{\rulebox}}            %
                      \addtolength{\ruledepth}{\rulevgap}                   %
                      \raisebox{0in}[\ruleheight][\ruledepth]               %
                               {\usebox{\rulebox}}}
\newcommand{\Axiom}[1]{\savebox{\rulebox}[\width][b]                        %
                               {$\frac{}{\raisebox{-0.03in}{$#1$}}$}        %
                      \settoheight{\ruleheight}{\usebox{\rulebox}}          %
                      \addtolength{\ruleheight}{\rulevgap}                  %
                      \settodepth{\ruledepth}{\usebox{\rulebox}}            %
                      \addtolength{\ruledepth}{\rulevgap}                   %
                      \raisebox{0in}[\ruleheight][\ruledepth]               %
                               {\usebox{\rulebox}}}
\newcommand{\RuleSide}[3]{\gap{\mbox{$#2$}} \hspace*{0.1in}               %
                            \Rule{#1}{#3}                          %
                          \hspace*{0.1in}\mbox{$#2$}}
\newcommand{\AxiomSide}[2]{\gap{\mbox{$#1$}} \hspace*{0.1in}              %
                             \Axiom{#2}                            %
                           \hspace*{0.1in}\mbox{$#1$}}
\newcommand{\RULE}[1]{\textbf{#1}}
\newcommand{\hg}{\quad}

\begin{document}

\title{Epigram 2\\
       more design thoughts}
\author{Conor McBride}
\maketitle

\section{Introduction}

With the underlying type theory beginning to settle, it's time to
think a bit more about how the high-level Epigram language might
work. This document contains some ideas, and delineates some gaps.

I shall write mainly in the present tense, intending that the
assertions I make become true at some point.

\section{The Coauthor Transducer Model}

As an experience, Epigram programming is interactive. As a document,
an Epigram program is the `minute' of an interaction. To load and
check an Epigram program is to replay the minuted interaction,
checking its plausibility.

Epigram operates as a \textbf{transducer}, consuming a source
file, and producing a modified source file, marked up with error
reports and responses to requests for help. The transducer is
idempotent. Epigram thus acts like a mechanical coauthor, offering
opinions and responding to requests for contribution.

Epigram has a high-level \textbf{source} language, in which
\texttt{.epi} documents are written, and a low-level \textbf{evidence}
language, into which documents are \textbf{elaborated}. The system
thus works by constructing a sequence of definitions in the evidence
language elaborating the definitions found in its input. The
elaboration process is, by its nature, incremental. The system
maintains the correspondence between regions of the source document
and definitions at the evidence layer: the output file may thus be
generated from the final \textbf{proof state} of the elaboration
process.

In addition to generating the output file, the system will (on
request) dump the final proof state in a textual format---a
\texttt{.gram} file. From a \texttt{.gram} file, both input and output
\texttt{.epi} files are recoverable, hence the patch from one to the
other is computable. Given a \texttt{.gram} file and
a patch to its output \texttt{.epi}, a new input \texttt{.epi} can be
constructed, which can be processed in turn, resulting in a new output
\texttt{.epi}, and hence a further patch. Editor integration is thus
unnecessary, but in any case easy.

A separate (or at least separable) component, based on Edwin Brady's
\textsc{epic} compiler, can generate an executable from a \texttt{.gram}
file. The former, if executed, offers a straightforward
read-eval-print loop for first-order expressions constructed over the
signature of definitions provided by the original source. Incomplete
definitions in the source may result in run-time errors: these are the
only run-time errors.  Such an executable, invoked with an argument,
treats that argument as a single command session.

Markdown is king.

In the event that Epigram programs ever become large enough to
necessitate modular development, \texttt{.gram} files will be generated
on a per-module basis. We may thus become interested in patches
between states of these \texttt{.gram} files as a means to transmit
changes downstream.

\section{Interactive Problem Solving}

Epigram documents show the declaration and partial solution of problems.
Problem solving is mediated as an interactive process. Every problem
in the source language can be encoded as a type in the evidence
language. Solving a problem in the source language amounts to finding
an inhabitant of the type which encodes it, by a process of
hierarchical refinement. This section is not supposed to be a definition:
it's an explorations.

A \textbf{document} is a series of \textbf{developments}. A
development comprises a \textbf{declaration} introducing a problem,
and a \textbf{refinement} reducing the problem to a hopefully
empty collection of subproblems. Too much abstraction: example, please!
\newcommand{\capbox}[2]{\raisebox{0.12in}{\begin{array}[t]{|l|}\hline
      #1 \vspace*{-0.05in}\\
    \hfill\mbox{\scriptsize{#2}} \\ \hline\end{array}}}
\newcommand{\Ts}[1]{\texttt{#1}\;}
\newcommand{\co}{\;:\;}
\newcommand{\sco}{;\;\;}
\newcommand{\cm}{,\;}
\[
\capbox{
  \capbox{
    \capbox{\Ts{x} \cm \Ts{y}\co \Ts{Nat}}{declaration}\\
    \Ts{--------------}\\
    \capbox{\Ts{x} \Ts{+} \Ts{y}}{template} \co \capbox{\Ts{Nat}}{type}
    }
    {declaration}\\ \\
  \capbox{
     \capbox{\Ts{x}\Ts{+}\Ts{y}}{problem}
       \capbox{\Ts{<=}\capbox{\Ts{induction}\Ts{x}}{eliminator}}{tactic}\\
     \capbox{
      \{\;\capbox{\capbox{\Ts{'zero}\Ts{+}\Ts{y}}{problem}
                  \capbox{\Ts{=} \capbox{\Ts{y}}{term}}{tactic}
                  \capbox{}{block}
          }{refinement}\\
      ;\;\capbox{\capbox{\Ts{'suc}\Ts{x'}\Ts{+}\Ts{y}}{problem}
                 \capbox{\Ts{=} \capbox{\Ts{'suc}
                   \capbox{\{.\;.\}}{shed}}{term}}{tactic}
                  \capbox{}{block}
         }{refinement}\\
      \}
      }{block}
    }
    {refinement}}
    {development}
\]

Lots of points show up from this example. I'll skim through them,
starting with the declaration: we give a template for addition,
hypothesizing declarations for the fields of the template, then giving
a type for the whole thing. This template determines the initial
problem. You can tell which tokens in the template are punctuation
rather than placeholders, because they don't have declarations!
Note that here, we can see value declarations, but we shall also have
propositional declarations, \(\texttt{:-}\;P\), where a proposition is
asserted anonymously---\(P\) then gives the shapes of problems for
refinement, i.e., propositional goals.

Refinements go `\textbf{problem} \textbf{tactic} \textbf{block}'. A
refinement is an offer to solve a problem of a particular form.
Correspondingly, the elaborator needs to check that the refinement
offered suits the problem or problems at hand: the rigidity of this
check will soften with time, but the minimum acceptable level is that
the refinement's problem freely determines the programmer's choice of
names for the variables in the problem. Working interactively, of course,
the problem parts of refinements are generated mechanically from the
problems at hand, ensuring a match.

A tactic provides the means to split a problem into zero or more
subproblems. The \texttt{<=} (`by') tactic invokes an \textbf{eliminator}
in the elimination-with-a-motive style. The \texttt{=} (`return') tactic
takes a term explaining how to compute the return value, and leaves no
subproblems. There will doubtless be others, not least the notorious
\texttt{with}.

The \textbf{shed}, \texttt{\{. .\}}, stands in place of a piece of program
(most syntactic categories, not just terms) to indicate that the programmer
has yet to choose something, thus causing elaboration to suspend. Note
that a tactic \texttt{= \{. .\}} does commit to return, where a
shed directly in a tactic place does not.

Now, in the coauthor transducer model, there is no trigger to fill in
a shed: one simply replaces the shed with code and reloads. Of course,
an editor could be customized to give this illusion, and the
patch-to-patch optimization would yield very similar results. What,
then, is the interaction?

The basic refinement interaction simply involves the system replacing
a null block with subproblems which are not accounted for.
Correspondingly,

\begin{alltt}
 x , y : Nat
-------------
 x + y : Nat

x + y <= induction x
\end{alltt}
%
gets fleshed out to
%
\begin{alltt}
let   x , y : Nat
     -------------
      x + y : Nat

x + y <= induction x
\{ 'zero + y    \{. .\}
; 'suc x' + y  \{. .\} 
\}
\end{alltt}

One might imagine going further, reporting coverage failure by inserting
subproblems into nonempty blocks. Such a circumstance might readily
arise if one added constructors to a datatype.

\subsection{Query and Response}

In addition to basic refinement functionality, we shall need to equip
programmers with the means to query the system and receive responses
in situ. Moreover, we shall need to equip the machine with the means
to insert remarks in a file. We need anchored comments. The machine
writes
\[
\Ts{\{!}\; \mathit{thing}\; \Ts{!!} \mathit{comment} \Ts{!\}}
\]
to indicate that the \(\mathit{comment}\) concerns the \(\mathit{thing}\).
Such a textual pattern is replaced by \(\mathit{thing}\) on reloading
(unless \(\mathit{thing}\) is blank, in which case the replacement is an
empty shed). Machine remarks are ephemeral responses, and should not be
put into the machine's mouth by the programmer. Meanwhile, an anchored query,
\[
\Ts{\{?}\; \mathit{thing}\; \Ts{??} \mathit{query} \Ts{?\}}
\]
means the same as \(\mathit{thing}\) (unless empty, in which case it means
the same thing as an empty shed) but attaches a query; on return, we can
expect a response in the form given above.

Your basic empty \texttt{\{? ?\}} means
``What goes here and what's in local scope?''. The basic use of
\(\Ts{\{!}\; \mathit{thing}\; \Ts{!!} \mathit{comment} \Ts{!\}}\)
is to report an error. A more amusing possibility would be to offer
a menu of choices for \(\mathit{thing}\).

It does not take a great leap of imagination to imagine how a fancier
interface might mediate point-and-click queries via this mechanism.
Again, one would need the patch-based transducer to support such a
rapid dialogue.

\subsection{Subproblems versus Side-Problems}

The hierarchical refinement of problems delivers the basic necessary
functionality, but it forces all sorts of auxiliary functions and lemmas
to be developed globally, in advance (spatially, if not temporally).
We might prefer to localize the development of helpers. I suggest that
we allow each refinement to carry a \texttt{where}-clause bearing a local
module of (partially) solved side-problems. Here's a simple example.

\begin{alltt}
 f : X -> Y; xs : List X
-------------------------
 map f xs : List Y

map f xs = hit xs where
\{  xs : List X
  -----------------
   hit xs : List X

  hit xs <= induction xs
  \{ hit 'nil = 'nil
  ; hit ('cons x xs') zs = 'cons (f x) (hit xs') \}\}
\end{alltt}

Scope-wise, the \texttt{where}-clause comes after the problem statement
but before tactic and block; textually, it can come before the tactic,
between tactic and block, or after block.

Note that on a good day with a following wind, the types of helper
functions can be inferred. One might hope to write
\begin{alltt}
 f : X -> Y; xs : List X
-------------------------
 map f xs : List Y

map f xs = \{? hit xs ?? helper ?\}
\end{alltt}
might deliver
\begin{alltt}
 f : X -> Y; xs : List X
-------------------------
 map f xs : List Y

map f xs = hit xs where
\{  xs : List X
  -----------------
   hit xs : List X

  hit xs \{! !\} \}
\end{alltt}

on the basis that the type \texttt{List X} is being pushed into
the helper application, comprising the function \texttt{hit} applied
(Miller-pattern-style) to distinct local variables (here just \texttt{xs})
whose types are known, thus determining \texttt{hit}'s declaration!
Where the helper function has need of implicit syntax, a little more
care will be needed to ensure control over what gets abstracted.

Propositional requirements arising in the course of programming might
also be discharged as side-problems in \texttt{where}-clauses. We do not
write proofs in terms: we rather expect proofs to be picked up from the
context, so side-problems allow us to seed the context appropriately.
Moreover, elaboration models type constraints as propositional equations,
so you can fix type-glitches (where provably equal types mismatch
definitionally) by adding the proofs to a \texttt{where}-clause.

\section{Kernel Mustered}

\newcommand{\SC}{\textsc}
\newcommand{\Set}[1]{\mathrm{Set}^{#1}}
\newcommand{\PI}[2]{(#1\!:\!#2)\to}
\newcommand{\SG}[2]{(#1\!:\!#2)\times}
\newcommand{\ONE}{\mathrm{1}}
\newcommand{\ifr}{\underline}
\newcommand{\lam}[1]{\lambda#1\mapsto}
\newcommand{\pr}[2]{[#1 | #2]}
\newcommand{\vd}{[]}
\newcommand{\cum}{\preceq}
\newcommand{\car}[1]{#1\:!}
\newcommand{\cdr}[1]{#1\:-}
\newcommand{\nm}{\Downarrow}
\newcommand{\SJ}[4]{#1\:\{#2 \mapsto #3\}\nm#4}
\newcommand{\letin}[3]{\mathrm{let}\: #1 = #2 \:\mathrm{in}\: #3}

It's a moving target, but not a bad plan to write it down once in a
while.

\[\begin{array}{rcll}
\SC{ChkTm} & ::= &  \Set{n} & n\in\mathbb{N} \\
  &|& \PI{v}{\SC{ChkTm}}\SC{ChkTm} & \mbox{dependent function type} \\
  &|& \SG{v}{\SC{ChkTm}}\SC{ChkTm} & \mbox{dependent pair type} \\
  &|& \ONE \\
  &|& \lam{v}\SC{ChkTm} \\
  &|& \pr{\SC{ChkTm}}{\SC{ChkTm}} & \mbox{pair}\\
  &|& \vd \\
  &|& \letin{v}{\SC{InfTm}}{\SC{ChkTm}} \\
  &|& \ifr{\SC{InfTm}} & \mbox{explicit direction change}
\medskip \\
\SC{InfTm} & ::= & v & \mbox{bound variable, untyped}\\
  &|& x_{:\SC{Nm}} & \mbox{free variable, typed}\\
  &|& f_{\mapsto\SC{Nm}:\SC{Nm}} & \mbox{definition, with value
    and type}\\
  &|& \SC{InfTm}\:\SC{ChkTm} & \mbox{application} \\
  &|& \car{\SC{InfTm}} & \mbox{car}\\
  &|& \cdr{\SC{InfTm}} & \mbox{cdr}\\
  &|& (:\SC{ChkTm}) & \mbox{identity cast at given type}
\end{array}\]

If you eliminate \(v\)'s with no binder,
definitions, defined variables, and identity casts, \(\SC{ChkTm}\)
becomes \(\SC{Nm}\) and \(\SC{InfTm}\) becomes \(\SC{Ne}\), the
\emph{normal} and \emph{neutral} terms, respectively. Let us define
untyped computation relations of the following forms:
\[\begin{array}{c}
\SJ{\SC{Nm}}{v}{\SC{Nm}}{\SC{Nm}}\\
\SJ{\SC{Ne}}{v}{\SC{Nm}}{\SC{Nm}}\\
\SC{Nm}\:\SC{Nm} \mapsto \SC{Nm}\\
\car{\SC{Nm}} \mapsto \SC{Nm}\\
\cdr{\SC{Nm}} \mapsto \SC{Nm}\\
\end{array}\]

Note that substituting a \SC{Ne} for a \(v\) preserves its \SC{Ne}-ness
and requires no worry about variable capture. Similarly, we may substitute
a \SC{ChkTm} for a \(v\) and get a \SC{ChkTm}.

The typing judgments are then as follows:
\[\begin{array}{lr}
(\{\Set{\omega}\}\cup\SC{Nm})\ni\SC{ChkTm}\nm\SC{Nm} & \mbox{type checking}\\
\SC{InfTm}\nm{\SC{Nm}}\in\SC{Nm} & \mbox{type inference}\\
\SC{Nm}\cum\SC{Nm} & \mbox{cumulativity}\\
\SC{Nm}\ni\SC{Nm}\equiv\SC{Nm} &
\mbox{equality} \\
\SC{Ne}\equiv\SC{Ne}\in\SC{Nm} & \mbox{equality} \\
\end{array}\]

Type checking. Where \(x\) or \(f\) appears in premises but not
conclusion, it is chosen fresh.
\[\begin{array}{c}
\Axiom{\Set{j}\ni\Set{i}\nm\Set{i}}\;j>i \quad
\Axiom{\Set{i}\ni\ONE\nm\ONE} \quad
\Rule{s\nm s'\in S\hg S\cum T}
  {T\ni\ifr{s}\nm s'}
\\

\Rule{\Set{i}\ni S\nm S' \hg \Set{i}\ni T[x_{:S}/v]\nm T'[x_{:S}/v]}
  {\Set{i}\ni\PI{v}{S}T\nm \PI{v}{S'}{T'}}
\quad
\Rule{\Set{i}\ni S\nm S' \hg \Set{i}\ni T[x_{:S}/v]\nm T'[x_{:S}/v]}
  {\Set{i}\ni\SG{v}{S}T\nm \SG{v}{S'}{T'}}
\\

\Rule{T[x_{:S}/u]\ni t[x_{:S}/v]\nm t'[x_{:S}/v]}
  {\PI{u}{S}{T}\ni\lam{v}t\nm\lam{v}t'}
\quad
\Rule{S\ni s\nm s'\hg \SJ{T}{v}{s'}{T'}\hg T' \ni t\nm t'}
  {\SG{v}{S}T \ni \pr{s}{t} \nm \pr{s'}{t'}}
\quad
\Axiom{\ONE\ni\vd\nm\vd}\\
\Rule{s\nm s'\in S\hg T\ni t[f_{\mapsto s':S}/v]\nm t'}
  {T\ni\letin{v}{s}{t}\nm t'}

\end{array}\]

Type inference. Note no rule for bound vars.
\[\begin{array}{c}
\Axiom{x_{:S}\nm\ifr{x_{:S}}\in S} \quad \Axiom{f_{\mapsto s : S}\nm s \in S}
\quad
\Rule{\Set{\omega}\ni T}
  {(:T)\nm\lam{v}v\in \PI{v}{T}T}
 \\
\Rule{f\nm f'\in\PI{v}{S}T \hg S\ni s\nm s'
   \hg \SJ{T}{v}{s'}{T'} \hg f'\:s'\mapsto t'}
  {f\:s\nm t' \in T'}
\\
\Rule{p\nm p'\in\SG{v}{S}T\hg \car{p'}\mapsto s'}
  {s' \in S}
\quad
\Rule{p\nm p'\in\SG{v}{S}T\hg \car{p'}\mapsto s' \hg \SJ{T}{v}{s'}{T'}
   \hg \cdr{p'}\mapsto t'}
  {\cdr{p}\nm t' \in T'}
\end{array}\]

Cumulativity.
\[\begin{array}{c}
\Axiom{\Set{i}\cum\Set{j}}\;i\le j
\quad
\Rule{S'\cum S \quad T[x_{S'}/u]\cum T'[x_{S'}/v]}
  {\PI{u}{S}T \cum \PI{v}{S'}T'}
\quad
\Rule{S\cum S' \quad T[x_{S}/u]\cum T'[x_{S}/v]}
  {\SG{u}{S}T \cum \SG{v}{S'}T'}
\quad
\Axiom{\ONE\cum\ONE}
\\
\Rule{\ifr{S} \equiv \ifr{T}\in\Set{n}}
  {S \cum T}
\end{array}\]

Equality for checkable terms.
\[\begin{array}{c}
\Axiom{\Set{j}\ni\Set{i}\equiv\Set{i}}\;j>i
\\
\Rule{\Set{i}\ni S'\equiv S \hg \Set{i}\ni T[x_{S'}/u] \equiv T'[x_{s'}/v]}
  {\Set{i}\ni \PI{u}{S}T\equiv\PI{v}{S'}{T'}}
\\
\Rule{\Set{i}\ni S\equiv S' \hg \Set{i}\ni T[x_{S}/u] \equiv T'[x_{s'}/v]}
  {\Set{i}\ni \SG{u}{S}T\equiv\SG{v}{S'}{T'}}
\\
\Rule{\PI{v}{S}T \ni f \hg \PI{v}{S}T \ni f' \hg
 f\:\ifr{x_{:S}}\mapsto t\hg f'\:\ifr{x_{:S}}\mapsto t' \hg T[x_{:S}/v]\ni t \equiv t'}
  {\PI{v}{S}T \ni f \equiv f'}
\\
\Rule{\begin{array}{c}
\SG{v}{S}T \ni p \hg \SG{v}{S}T \ni p' \\
\car{p}\mapsto s \hg \cdr{p}\mapsto t \hg 
\car{p'}\mapsto s' \hg \cdr{p'}\mapsto t' \\
S\ni s\equiv s' \hg \SJ{T}{v}{s}{T'} \hg T'\ni t\equiv t'
\end{array}}
  {\SG{v}{S}T \ni p \equiv p'}
\quad
\Rule{\ONE\ni t \hg \ONE \ni t'}
  {\ONE\ni t \equiv t'}
\\
\Rule{s\equiv t \in S \hg S\cum T}
  {T\ni\ifr{s}\equiv\ifr{t}}
\end{array}\]

Equality for inferrable terms.
\[\begin{array}{c}
\Axiom{x_{:S}\equiv x_{:S}\in S}
\quad
\Rule{f\equiv f'\in\PI{v}{S}T \hg S\ni s\equiv s' \hg \SJ{T}{v}{s}{T'}}
  {f\:s\equiv f'\:s'\in T'}
\\
\Rule{p\equiv p'\in\SG{v}{S}T}
  {\car{p}\equiv \car{p'}\in S}
\quad
\Rule{p\equiv p'\in\SG{v}{S}T}
  {\cdr{p}\equiv \cdr{p'}\in T[\car{p}/v]}
\end{array}\]

Computation.
\[\begin{array}{c}
\Rule{\SJ{t}{v}{s}{t'}}
  {(\lam{v}t)\:s\mapsto t'} \quad
\Axiom{\ifr{f}\:s \mapsto \ifr{f\:s}}
\quad
\Axiom{\car{\pr{s}{t}}\mapsto s}\quad
\Axiom{\car{\ifr{p}} \mapsto \ifr{\car{p}}}\quad
\Axiom{\cdr{\pr{s}{t}}\mapsto t}\quad
\Axiom{\cdr{\ifr{p}} \mapsto \ifr{\cdr{p}}}
\end{array}\]

Normal substitution.
\[\begin{array}{c}
\Axiom{\SJ{\Set{i}}{v}{r}{\Set{i}}}
\\
\Rule{\SJ{S}{v}{r}{S'}\quad\SJ{T[x_{S'}/u]}{v}{r}{T'[x_{S'}/u]}}
  {\SJ{(\PI{u}{S}T)}{v}{r}{\PI{u}{S'}T'}}
\\
\Rule{\SJ{S}{v}{r}{S'}\quad\SJ{T[x_{S'}/u]}{v}{r}{T'[x_{S'}/u]}}
  {\SJ{(\SG{u}{S}T)}{v}{r}{\SG{u}{S'}T'}}
\\
\Rule{\SJ{t[x_{S'}/u]}{v}{r}{t'[x_{S'}/u]}}
  {\SJ{(\lam{u}{t})}{v}{r}{\lam{u}t'}}
\quad
\Rule{\SJ{s}{v}{r}{s'}\hg\SJ{t}{v}{r}{t'}}
  {\SJ{\pr{s}{t}}{v}{r}{\pr{s'}{t'}}}
\quad
\Axiom{\SJ{\vd}{v}{r}{\vd}}
\\
\Rule{\SJ{t}{v}{r}{t'}}
  {\SJ{\ifr{t}}{v}{r}{t'}}
\end{array}\]

Neutral substitution.
\[\begin{array}{c}
\Axiom{\SJ{v}{v}{r}{r}} \quad
\Axiom{\SJ{x_{:S}}{v}{r}{\ifr{x_{:S}}}} \\
\Rule{\SJ{f}{v}{r}{f'}\hg\SJ{s}{v}{r}{s'}\hg f'\:s'\mapsto t'}
  {\SJ{(f\:s)}{v}{r}{t'}} \\
\Rule{\SJ{p}{v}{r}{p'}\hg\car{p'}\mapsto s'}
  {\SJ{(\car{p})}{v}{r}{s'}} \quad
\Rule{\SJ{p}{v}{r}{p'}\hg\cdr{p'}\mapsto t'}
  {\SJ{(\cdr{p})}{v}{r}{t'}}
\end{array}\]

\end{document}